---
title: "agriPairMetaFlow 0.3.0: From Agronomic Question to Dependence-aware Meta-regression and Dose Response"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{agriPairMetaFlow 0.3.0: From Agronomic Question to Dependence-aware Meta-regression and Dose Response}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(agriPairMetaFlow)
```


# 1. Why the package starts with design

A treatment-control meta-analysis is not defined only by two columns of means. The experimental unit, reused controls, true matching, response scale, and uncertainty determine what can be synthesized. Version 0.3.0 connects auditing, effect-size construction, dependence-aware fitting, robust inference, prespecified meta-regression, quantitative moderator curves, and correlated dose-response synthesis.

# 2. Inspect and audit

```{r}
apm_validate(maize_n_shared, "summary", strict=FALSE)
apm_audit(maize_n_shared, "full")
apm_plan(maize_n_shared, study=study_id, experiment=experiment_id, treatment=treatment, control=control, response=mean_t, dose=N_rate)
```

The audit warns that the same control is reused. This is scientifically consequential. From version 0.2.0 onward the covariance can be represented explicitly with `apm_vcov()` and fitted with multilevel or meta-regression models rather than treating these contrasts as independent.

# 3. Effect sizes

```{r}
e_rr <- apm_effect_size(covercrop_variability,"lnRR",m_t=mean_t,sd_t=sd_t,n_t=n_t,m_c=mean_c,sd_c=sd_c,n_c=n_c)
e_vr <- apm_effect_size(covercrop_variability,"VR",m_t=mean_t,sd_t=sd_t,n_t=n_t,m_c=mean_c,sd_c=sd_c,n_c=n_c)
e_cvr <- apm_effect_size(covercrop_variability,"CVR",m_t=mean_t,sd_t=sd_t,n_t=n_t,m_c=mean_c,sd_c=sd_c,n_c=n_c)
```

lnRR addresses mean response; VR and CVR ask different questions about variability. A treatment may increase the expected yield yet decrease relative variability, which can be agronomically valuable.

# 4. Core model, heterogeneity, prediction, and thresholds

```{r}
fit <- apm_fit(agri_effects_benchmark)
apm_heterogeneity(fit, ci=FALSE)
apm_prediction(fit, transform="percent")
apm_threshold(fit, 5, scale="percent")
```

A confidence interval describes uncertainty about the average effect. A prediction interval addresses the range expected for a future context under the fitted random-effects model. They answer different questions and should not be interchanged.

# 5. Subgroups and visualization

```{r}
apm_subgroup(agri_effects_benchmark, crop, test="z")
apm_forest(fit, transform="percent", columns=c("crop","dose"))
apm_funnel(fit, contour=TRUE)
apm_table(fit,"model",transform="percent")
```

A funnel plot is a diagnostic display, not a publication-bias verdict. Subgroup-specific significance is likewise not evidence that subgroup effects differ; the between-subgroup test is the relevant comparison.


# 6. Dependence-aware meta-regression

```{r}
mz_es <- apm_effect_size(maize_n_shared,"lnRR",m_t=mean_t,sd_t=sd_t,n_t=n_t,m_c=mean_c,sd_c=sd_c,n_c=n_c)
mz_V <- apm_vcov(mz_es,cluster=experiment_id,shared_control=TRUE)
mz_reg <- apm_metareg(mz_es,~N_rate+soil_texture,V=mz_V,random=~1|study_id/effect_id)
apm_marginal_effects(mz_reg,variables="soil_texture",transform="percent")
```

The moderator analysis retains the sampling covariance induced by the common zero-N arm.

# 7. Quantitative moderator curves

```{r}
irr_es <- apm_effect_size(irrigation_climate,"lnRR",m_t=mean_t,sd_t=sd_t,n_t=n_t,m_c=mean_c,sd_c=sd_c,n_c=n_c)
irr_curve <- apm_metareg_curve(irr_es,rainfall,"quadratic")
apm_curve_features(irr_curve,c("slope","turning_point"))
apm_predict_context(irr_curve,data.frame(rainfall=c(700,1000)),transform="percent")
```

These are across-study moderator relationships. A curve feature is not automatically a causal agronomic optimum.

# 8. Correlated dose-response

```{r}
dose_fit <- apm_dose_response(fertilizer_dose_response,effect=lnRR,dose=N_rate,study=study_id,form="quadratic",method="fixed")
apm_dose_plot(dose_fit,transform="percent")
```

Dose-response is kept conceptually separate because several doses within one experiment share the same reference group.

# Version 0.5.0: diagnosing, stress-testing, and communicating the synthesis

A fitted model is not the end of the workflow. Before communicating an agronomic synthesis, ask whether the conclusion is unusually dependent on one independent study, whether subsets reveal unresolved heterogeneity structure, whether small-study patterns change under explicit selection assumptions, and whether the final figure distinguishes mean-effect uncertainty from prediction for a new setting.

```{r}
fit05 <- apm_fit(agri_effects_benchmark)
loo05 <- apm_leave_one_out(fit05, unit = "study")
inf05 <- apm_influence(fit05, unit = "study", plot = FALSE)
bias05 <- apm_bias(fit05, methods = c("egger", "rank", "trimfill"))
apm_orchard(fit05, transform = "percent")
apm_explain(fit05, audience = "scientific", transform = "percent")
```

These functions answer different questions. Leave-one-study-out sensitivity asks whether the pooled result or heterogeneity depends strongly on one independent evidence unit. Influence metrics locate unusual leverage or impact. GOSH examines many subsets and is useful for discovering unresolved structures. Funnel-asymmetry and selection analyses probe possible small-study or publication-selection mechanisms under specific assumptions. None of these procedures provides an automatic study-exclusion rule.

When many moderators are available, `apm_moderator_screen()` can be used as an exploratory MetaForest stage. It should be followed by scientifically specified meta-regression rather than interpreted as a confirmatory variable-selection procedure. The package keeps this distinction in the returned class and caution text.

For reporting, `apm_report()` summarizes an already fitted object and records provenance; it does not refit the model. `apm_export()` preserves the complete object in RDS or creates numeric tables and publication-quality figures. Large GOSH, MetaForest tuning, wild bootstrap, and Bayesian computation should be precomputed in the developer workflow when they would make routine vignette rendering expensive.
