---
title: "Data, Uncertainty, Effect Sizes, and True Pairing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data, Uncertainty, Effect Sizes, and True Pairing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(agriPairMetaFlow)
```


# Data import and validation

```{r}
f <- system.file("extdata","agri_uncertainty_mixed.csv",package="agriPairMetaFlow")
x <- apm_read(f)
apm_validate(x, "summary", strict=FALSE)
apm_audit(x,"full")
```

# Recovering reported uncertainty

```{r}
u1 <- apm_recover_uncertainty(agri_uncertainty_mixed,mean=mean_yield,n=n,se=se_yield,method="se")
u2 <- apm_recover_uncertainty(agri_uncertainty_mixed,mean=mean_yield,n=n,cv=cv_percent,method="cv")
u3 <- apm_recover_uncertainty(agri_uncertainty_mixed,mean=mean_yield,n=n,mse=residual_mse,method="mse")
```

The provenance column distinguishes reported values from algebraically reconstructed values. Missing uncertainty is never silently guessed.

# Independent and paired effects

```{r}
p1 <- apm_plan(wheat_paired_blocks,study=study_id,treatment=treatment,control=control,response=mean_t,block=block_id,design="paired")
e1 <- apm_effect_size(wheat_paired_blocks,"lnRR",design="paired",m_t=mean_t,sd_t=sd_t,n_t=n_pairs,m_c=mean_c,sd_c=sd_c,n_c=n_pairs,r=r_tc)
e2 <- apm_effect_size(pest_suppression_binary,"RR",event_t=event_t,event_c=event_c,n_t=n_t,n_c=n_c)
e3 <- apm_effect_size(covercrop_variability,"CVR",m_t=mean_t,sd_t=sd_t,n_t=n_t,m_c=mean_c,sd_c=sd_c,n_c=n_c)
```

Pairing is a property of the primary design, not of the existence of a treatment and a control. The paired route therefore requires an explicit correlation for continuous summary effects.

# Tables

```{r}
apm_table(e1,"effects")
apm_table(e2,"effects")
apm_table(e3,"effects")
```
