---
title: "Shared controls, paired designs, and dependence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shared controls, paired designs, and dependence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
library(agriPairMetaFlow)
```

## Why dependence matters

Agronomic experiments commonly compare several fertilizer rates, irrigation regimes, or management treatments with the same untreated control. The resulting effect sizes reuse information from the control arm and therefore do not have independent sampling errors. Treating them as independent artificially increases the apparent amount of information.

## Detect shared controls before fitting a model

```{r}
sc <- apm_shared_control(
  maize_n_shared,
  study = experiment_id,
  control_id = control,
  treatment_id = treatment
)
sc
```

The audit is intentionally separated from model fitting. Conflicting duplicated control means, standard deviations, or sample sizes are flagged rather than silently reconciled.

## Build effect sizes and the sampling covariance matrix

```{r}
es <- apm_effect_size(
  maize_n_shared, "lnRR",
  m_t=mean_t, sd_t=sd_t, n_t=n_t,
  m_c=mean_c, sd_c=sd_c, n_c=n_c
)
V <- apm_vcov(es, cluster=experiment_id, shared_control=TRUE)
V[1:6,1:6]
```

`apm_vcov()` delegates the general covariance construction to `metafor::vcalc()`. When treatment and control identifiers are available, they are passed as the two contrast groups; arm sample sizes are supplied as weights when possible. The object records the assumptions used and checks symmetry, the diagonal, and positive semidefiniteness.

## Paired does not merely mean treatment versus control

A genuine paired design uses matched or repeated experimental units. The `wheat_paired_blocks` data contain an explicit treatment-control correlation used when computing paired effect sizes.

```{r}
pes <- apm_effect_size(
  wheat_paired_blocks, "lnRR", design="paired",
  m_t=mean_t, sd_t=sd_t, n_t=n_pairs,
  m_c=mean_c, sd_c=sd_c, n_c=n_pairs,
  r=r_tc
)
```

When several effects belong to the same dependency block, `apm_pair_vcov()` can represent a compound correlation assumption. Unknown correlations should be examined by sensitivity analysis rather than chosen to obtain a preferred result.

## Sensitivity to an unknown correlation

```{r}
sm <- apm_effect_size(
  soil_management_multiresponse, "lnRR",
  m_t=mean_t, sd_t=sd_t, n_t=n_t,
  m_c=mean_c, sd_c=sd_c, n_c=n_c
)
rho_check <- apm_rho_sensitivity(
  sm,
  rho=c(.25,.50,.75),
  build_vcov=list(cluster="study_id", type="outcome")
)
rho_check
```

A stable conclusion should not depend critically on an arbitrary, weakly justified within-study correlation. The complete grid is retained, including failed fits.
