---
title: "Multilevel and robust inference for dependent agronomic evidence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multilevel and robust inference for dependent agronomic evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
library(agriPairMetaFlow)
```

## Hierarchical evidence structures

A paper may contain experiments, sites, years, outcomes, and several treatment-control contrasts. `apm_multilevel()` exposes this hierarchy through an agronomic interface and fits the corresponding model with `metafor::rma.mv()`.

```{r}
es <- apm_effect_size(
  maize_n_shared, "lnRR",
  m_t=mean_t, sd_t=sd_t, n_t=n_t,
  m_c=mean_c, sd_c=sd_c, n_c=n_c
)
V <- apm_vcov(es, cluster=experiment_id)
ml <- apm_multilevel(
  es,
  random=~1|study_id/effect_id,
  V=V,
  test="z"
)
ml
```

## Where does heterogeneity occur?

```{r}
apm_variance_components(ml)
```

The reported percentages describe allocation of the modeled heterogeneity components. They are not percentages of total observed variance and should not be interpreted as such.

## Cluster-robust inference

For meta-regression or dependent effect structures, model-based standard errors can be complemented by cluster-robust inference. The package defaults to CR2 because the bias-reduced adjustment is designed for improved small-sample performance.

```{r, eval=requireNamespace("clubSandwich", quietly=TRUE)}
irr <- apm_effect_size(
  irrigation_climate, "lnRR",
  m_t=mean_t, sd_t=sd_t, n_t=n_t,
  m_c=mean_c, sd_c=sd_c, n_c=n_c
)
fit <- apm_fit(irr, mods=~rainfall)
cr2 <- apm_robust(fit, cluster=study_id)
cr2
apm_compare_inference(fit, robust=cr2)
```

Low denominator degrees of freedom are reported because a nominally large number of effect sizes does not guarantee precise robust inference when the number or leverage distribution of independent clusters is unfavorable.

## Cluster wild bootstrap

The wild-bootstrap layer is optional and intentionally does not run heavy resampling during ordinary installation.

```{r, eval=FALSE}
wb <- apm_wild_bootstrap(
  fit,
  cluster=study_id,
  constraints=2,
  R=9999,
  seed=20260826
)
apm_compare_inference(fit, robust=cr2, wild=wb)
```

The result stores the bootstrap distribution settings, seed, number of independent clusters, p-value, and Monte Carlo standard error. Conventional, CR2, and bootstrap inference are displayed together; the package does not choose whichever method gives the smallest p-value.
