---
title: "Multivariate treatment-control meta-analysis in agronomy"
author: "agriPairMetaFlow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multivariate treatment-control meta-analysis in agronomy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
library(agriPairMetaFlow)
```

# Why multivariate synthesis matters

Agronomic experiments rarely measure only one endpoint. A soil-management study may report yield, soil organic carbon, infiltration, nutrient-use efficiency, and greenhouse-gas emissions from the same experimental units. Separate univariate meta-analyses are useful for endpoint-specific questions, but they discard the correlation created by measuring several responses in the same study and cannot estimate a between-outcome heterogeneity covariance structure.

`agriPairMetaFlow` therefore treats multivariate synthesis as a distinct estimand and data problem. A study identifier must identify the independent experimental unit contributing the response vector. The `outcome` variable identifies endpoints, while `yi` and `vi` remain the effect estimate and its sampling variance on a coherent scale.

# Teaching data

`soil_management_multiresponse` contains three outcomes per synthetic experiment. The response-ratio effects are already available as `yi`; the package deliberately labels the data as synthetic teaching data rather than field evidence.

```{r data}
head(soil_management_multiresponse)
with(soil_management_multiresponse, table(mv_study_id, outcome))
```

The second example, `biochar_multiresponse`, contains yield, soil carbon, and N2O responses from eight synthetic studies.

```{r biochar}
head(biochar_multiresponse)
with(biochar_multiresponse, table(study_id, outcome))
```

# Start with the sampling covariance

If the within-study covariance between outcomes is known or can be reconstructed, it should be supplied. The simplest diagonal matrix assumes zero within-study outcome correlation and is useful as a baseline, not as a default scientific conclusion.

```{r multivariate-baseline}
V0 <- diag(soil_management_multiresponse$vi)
mv0 <- apm_multivariate(
  soil_management_multiresponse,
  outcome = outcome,
  study = mv_study_id,
  V = V0,
  structure = "UN"
)
mv0
summary(mv0)
```

An unstructured between-outcome covariance is flexible but parameter intensive. A compound-symmetry or diagonal structure can be evaluated when scientifically justified.

```{r multivariate-structures}
mv_cs <- apm_multivariate(
  soil_management_multiresponse,
  outcome = outcome,
  study = mv_study_id,
  V = V0,
  structure = "CS"
)
mv_diag <- apm_multivariate(
  soil_management_multiresponse,
  outcome = outcome,
  study = mv_study_id,
  V = V0,
  structure = "DIAG"
)
```

# Moderators in a multivariate model

Moderators are expanded by outcome. This allows the treatment-control effect to vary with climate while retaining outcome-specific intercepts.

```{r multivariate-moderator}
mv_climate <- apm_multivariate(
  soil_management_multiresponse,
  outcome = outcome,
  study = mv_study_id,
  V = V0,
  mods = ~ climate_zone,
  structure = "UN"
)
```

The modifier should be prespecified and interpreted as an association in the meta-regression. It is not automatically a causal effect of climate.

# Unknown within-study correlations

Published agronomic papers often report marginal standard errors but omit cross-outcome covariances. A single guessed correlation should not be hidden inside the analysis. `apm_mvcor_sensitivity()` reconstructs block covariance matrices over a prespecified correlation grid and refits the same multivariate model.

```{r rho-sensitivity}
sens1 <- apm_mvcor_sensitivity(
  soil_management_multiresponse,
  outcome = outcome,
  study = mv_study_id,
  rho = c(0, .25, .50, .75)
)
sens1
```

The same sensitivity workflow can be used for the biochar example.

```{r rho-biochar}
sens2 <- apm_mvcor_sensitivity(
  biochar_multiresponse,
  outcome = outcome,
  study = study_id,
  rho = seq(0, .8, by=.2),
  metric = "tau"
)
```

A moderator can be retained across the entire sensitivity grid.

```{r rho-moderator}
sens3 <- apm_mvcor_sensitivity(
  soil_management_multiresponse,
  outcome = outcome,
  study = mv_study_id,
  rho = c(0, .3, .6),
  fit_args = list(mods = ~ climate_zone)
)
```

The scientific question is not which value of rho gives the most convenient result. The question is whether the substantive conclusion changes across a plausible set of correlations.

# Visualization

The outcome forest emphasizes endpoint-specific pooled effects.

```{r mvplots}
apm_multivariate_plot(mv0, type="outcome_forest", transform="percent")
```

When a between-outcome correlation matrix is estimable, its structure can be inspected directly.

```{r mvcorr}
if (!is.null(mv0$between_cor)) {
  apm_multivariate_plot(mv0, type="correlation")
}
```

A third presentation is available for the outcome summary on the model scale.

```{r mvpred}
apm_multivariate_plot(mv0, type="prediction", transform="none")
```

# Backend cross-check

The default multivariate backend is `metafor::rma.mv()`. `mixmeta` is optional and useful as an independent implementation and for alternative multivariate structures. The 0.4.0 adapter requires a complete outcome profile for every study when `backend="mixmeta"`.

```{r mixmeta, eval=FALSE}
mv_mix <- apm_multivariate(
  biochar_multiresponse,
  outcome = outcome,
  study = study_id,
  V = diag(biochar_multiresponse$vi),
  structure = "CS",
  backend = "mixmeta"
)
```

During local validation, estimates and covariance matrices should be compared directly with independently specified `metafor` and `mixmeta` fits. Agreement is a numerical validation gate, not merely a documentation example.

# Interpretation checklist

A multivariate result should state the effect scale, the independent study unit, the origin of the sampling covariance matrix, the assumed or estimated correlation structure, the between-outcome heterogeneity structure, the number of studies and outcome profiles, and the sensitivity of conclusions to unknown correlations. Missing covariance information should never be described as if it were observed.
