---
title: "Bayesian meta-analysis, prediction, and agronomic relevance"
author: "agriPairMetaFlow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian meta-analysis, prediction, and agronomic relevance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
library(agriPairMetaFlow)
```

# Bayesian synthesis is an estimand choice, not a decoration

The Bayesian workflow in `agriPairMetaFlow` is designed to make priors, posterior targets, prediction, and computational diagnostics explicit. It does not define a treatment as useful merely because a posterior probability crosses a package-defined cutoff.

The 0.4.0 layer supports three complementary backends. `bayesmeta` provides deterministic numerical integration for normal-normal meta-analysis and meta-regression. `RoBMA` provides JAGS-based Bayesian models and model averaging, including publication-bias components. `brms` provides a flexible Stan-based route for multilevel structures. Heavy Bayesian packages remain optional dependencies.

# 1. Declare priors before fitting

A prior is an inspectable object.

```{r priors}
p1 <- apm_prior(
  effect = list(dist="normal", mean=0, sd=.20),
  tau = list(dist="halfnormal", scale=.20),
  notes = "Weakly informative prior on the log response-ratio scale"
)
p1
```

A moderator prior can be declared separately.

```{r moderator-prior}
p2 <- apm_prior(
  effect = list(dist="normal", mean=0, sd=.20),
  tau = list(dist="halfnormal", scale=.15),
  moderators = list(rainfall=list(dist="normal", mean=0, sd=.001))
)
```

Model-probability declarations are stored rather than silently renormalized.

```{r probability-prior}
p3 <- apm_prior(
  effect=list(dist="normal", mean=0, sd=.30),
  model_probability=list(effect=.5, heterogeneity=.5)
)
```

# 2. Inspect implications before seeing the posterior

For lnRR, a Normal prior centered at zero can be translated to ratio and percentage scales.

```{r prior-check}
pc1 <- apm_prior_check(p1, measure="lnRR", thresholds=log(c(.90,1.10)), draws=5000, seed=101)
pc1
```

For a quantitative moderator, plausible values can be supplied before model fitting.

```{r prior-context}
pc2 <- apm_prior_check(
  p2,
  measure="lnRR",
  x=data.frame(rainfall=c(600,900,1200)),
  draws=5000,
  seed=101,
  plot=FALSE
)
```

The same operation is meaningful for variability ratios.

```{r prior-cvr}
pc3 <- apm_prior_check(
  apm_prior(effect=list(dist="normal",mean=0,sd=.20)),
  measure="CVR",
  draws=5000,
  seed=101,
  plot=FALSE
)
```

# 3. Fit the normal-normal model

The examples below use optional Bayesian backends and are intentionally not evaluated in lightweight vignette builds. They are mandatory during the package's local validation protocol.

```{r bayesmeta-fit, eval=FALSE}
b1 <- apm_bayes(
  agri_effects_benchmark,
  prior=p1,
  backend="bayesmeta"
)
b1
```

A Bayesian meta-regression uses `bayesmeta::bmr()` through the same API.

```{r bayesmeta-reg, eval=FALSE}
es_irrig <- apm_effect_size(
  irrigation_climate, "lnRR",
  m_t=mean_t, sd_t=sd_t, n_t=n_t,
  m_c=mean_c, sd_c=sd_c, n_c=n_c
)
es_irrig$rainfall <- irrigation_climate$rainfall
b2 <- apm_bayes(es_irrig, mods=~rainfall, prior=p2, backend="bayesmeta")
```

A robust/model-averaged backend can be requested explicitly when installed.

```{r robma-fit, eval=FALSE}
b3 <- apm_bayes(
  agri_effects_benchmark,
  backend="RoBMA",
  bias_adjust=TRUE,
  seed=2026
)
```

# 4. Distinguish mean effects from prediction

Posterior uncertainty in the pooled mean is not the same as uncertainty in the true effect of a new study.

```{r bayes-predict, eval=FALSE}
mean_post <- apm_bayes_predict(b1, predictive=FALSE, transform="percent")
new_study <- apm_bayes_predict(b1, predictive=TRUE, transform="percent")
mean_post
new_study
```

For meta-regression, prediction is conditional on a stated agronomic context.

```{r bayes-context, eval=FALSE}
pred_rain <- apm_bayes_predict(
  b2,
  newdata=data.frame(rainfall=c(700,900,1100)),
  predictive=TRUE,
  transform="percent"
)
```

For `RoBMA`, the package uses the backend's `type="estimate"` target for a latent true-effect prediction. `type="response"` would add the future sampling error and answers a different question.

# 5. Agronomic relevance thresholds

Suppose a researcher defines a 5% yield increase as the smallest benefit that merits agronomic attention. The probability is then computed relative to that declared threshold.

```{r bayes-threshold, eval=FALSE}
t1 <- apm_bayes_threshold(b1, threshold=5, scale="percent", direction="greater")
t2 <- apm_bayes_threshold(b1, threshold=5, scale="percent", direction="greater", predictive=TRUE)
t3 <- apm_bayes_threshold(b1, threshold=0, scale="model", rope=c(-.05,.05))
```

`t1` concerns the pooled effect. `t2` concerns a new true study effect and is normally more diffuse. `t3` illustrates a region of practical equivalence. None of these operations imposes a universal decision threshold.

# 6. Diagnostics are backend specific

`bayesmeta` uses deterministic DIRECT integration, so MCMC diagnostics such as R-hat and effective sample size are not defined and must not be fabricated.

```{r diagnostics, eval=FALSE}
d1 <- apm_bayes_diagnostics(b1, plot=FALSE)
d2 <- apm_bayes_diagnostics(b2, checks=c("convergence","ppc"), plot=FALSE)
d3 <- apm_bayes_diagnostics(b3, checks=c("convergence","ess","mcse","ppc"), plot=TRUE)
```

For MCMC backends, convergence is a gate to substantive interpretation. Local validation must inspect R-hat/ESS/MCSE and backend-specific warnings before results are reported.

# 7. Model comparison

Predictive criteria and posterior model probabilities are different quantities. They should not be converted into p-values or used to automate model selection.

```{r comparison, eval=FALSE}
apm_bayes_compare(brms_linear, brms_quadratic, criterion="loo")
apm_bayes_compare(brms_linear, brms_quadratic, criterion="waic")
apm_bayes_compare(robma_ensemble, criterion="model_probability")
```

# 8. Reporting

A Bayesian agronomic meta-analysis should report the effect measure and natural-scale interpretation, all prior distributions and their scientific rationale, prior-predictive checks, posterior pooled effects, prediction for a new true study effect, heterogeneity, agronomic threshold probabilities only when the threshold is externally justified, convergence diagnostics, sensitivity to reasonable prior alternatives, backend/version/seed, and any publication-bias model as a sensitivity analysis rather than proof that bias exists.
