---
title: "agriPairMetaFlow 1.0: Integrated Workflow, Capabilities, and Release Readiness"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Integrated workflow and release readiness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
library(agriPairMetaFlow)
```

# 1. Why version 1.0 adds integration rather than another estimator

The stable 1.0 layer does not introduce a new meta-analytic estimator. Earlier
releases already cover effect sizes, dependence, multilevel and robust
inference, quantitative moderators, dose-response, multivariate outcomes,
Bayesian sensitivity, diagnostic analysis, publication-bias sensitivity,
visualization, reporting, and export. Version 1.0 connects those capabilities
without concealing scientific decisions.

`apm_workflow()` therefore retains every intermediate object. Its
`routing_log` states whether a decision was supplied by the analyst, inferred
from explicit design information, or derived from another package object.
Automatic routing is a convenience, not a substitute for understanding the
experimental design.

# 2. Three agronomic integrated workflows

A maize nitrogen synthesis with several doses sharing the same zero-N control
must not treat the contrasts as independent. The optional CR2 step is executed
only when its backend is installed.

```{r workflow-shared, eval=requireNamespace("clubSandwich", quietly=TRUE)}
w_maize <- apm_workflow(
  maize_n_shared,
  measure="lnRR",
  dependence="shared_control",
  model="multilevel",
  robust=TRUE
)
w_maize$routing_log
```

A climatic moderator analysis illustrates automatic random-effects routing and
a five-percent agronomic relevance threshold.

```{r workflow-irrigation}
w_irrig <- apm_workflow(
  irrigation_climate,
  measure="lnRR",
  moderators=~rainfall+mean_temp,
  threshold=5
)
w_irrig$routing_log
```

Paired analysis is only selected when paired sample-size/correlation information
is explicit.

```{r workflow-paired}
w_pair <- apm_workflow(
  wheat_paired_blocks,
  measure="lnRR",
  dependence="paired",
  model="random"
)
w_pair$routing_log
```

# 3. Capability registry

The registry separates the stable public feature from the package that happens
to implement a specialized calculation. This allows an optional backend to be
updated or replaced without changing the scientific meaning of the user-facing
workflow.

```{r capabilities-all}
apm_capabilities()
```

```{r capabilities-bayes}
apm_capabilities("bayesian", detail="full")
```

```{r capabilities-dose}
apm_capabilities("dose-response", installed=TRUE)
```

An unavailable optional backend is not a failure of the core package. It means
that the corresponding optional capability cannot be validated or used until
the dependency is installed.

# 4. Non-mutating installation doctor

The doctor does not install R packages, JAGS, CmdStan, Pandoc, Quarto, or system
libraries. It only reports what is present and how the result should be
interpreted.

```{r doctor-core}
apm_doctor()
```

```{r doctor-full}
apm_doctor(full=TRUE, check_backends=TRUE)
```

The deepest check is appropriate on the release-validation computer. It remains
a diagnostic convenience and is not a substitute for `testthat`, numerical
golden tests, vignette rendering, `R CMD build`, or `R CMD check --as-cran`.

```{r doctor-release, eval=FALSE}
apm_doctor(full=TRUE, check_render=TRUE, check_examples=TRUE)
```

# 5. Reading the routing log

For every integrated analysis inspect at least:

1. the detected or declared dependence structure;
2. the effect-size design and backend measure;
3. whether a full sampling covariance matrix was created;
4. the selected model family;
5. moderator and relevance-threshold decisions;
6. any robust or Bayesian sensitivity route.

The workflow is intentionally reversible. Every component can be recomputed
manually using the lower-level function recorded by the route. During local
validation, the manual and orchestrated results must agree numerically within
the tolerance appropriate to the deterministic backend.

# 6. Release principle

Static source auditing is useful for catching missing files, documentation
coverage, obvious source inconsistencies, and packaging divergence. It is not
runtime validation. The formal 1.0 release exists only after local R validation
has regenerated documentation, executed tests and examples, rendered normal
vignettes, built a source tarball, frozen that tarball, and run
`R CMD check --as-cran` on that exact immutable artifact.
